---
title: "BME4450_Assignment8"
author: "Kelsie Reinaltt"
date: "10/31/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rafalib)
library(RColorBrewer)
library(GSE5859Subset)
data(GSE5859Subset)
```

# Chapter: Machine Learning

# Machine Learning, Exersices on Pages 374-375
## Question 1
#### B) 19 & 14 are the furthest away from each other
```{r}
set.seed(1)
m = 10000
n = 24
x = matrix(rnorm(m*n),m,n)
colnames(x)=1:n
head(x)
hc = hclust( dist( t(x)))
plot(hc)
```

## Question 2
#### The standard error  is 0.8986657
```{r}
set.seed(1)
m = 10000
n = 24
nc = replicate(100,{
x = matrix(rnorm(m*n),m,n)
hc = hclust( dist( t(x)))
length(unique(cutree(hc,h=143)))
})
plot(table(nc)) ## look at the distribution
popsd(nc)
```

## Question 3
####  A) Date is driving the clusters
```{r}
mds=cmdscale(dist(t(geneExpression)))
set.seed(10)
result=kmeans(t(geneExpression),5)
mypar(1,1)
plot(mds,bg=result$cl,pch=21)
table(sampleInfo$date,result$cluster)[,c(4,1,5,3,2)]
```

## Question 4
#### A) group of chrY genes are higher in group 0 and appear to drive the clustering. Within those clusters there appears to be clustering by 
```{r}
##load libraries
library(rafalib)
library(gplots)
library(matrixStats)
library(RColorBrewer)
##make colors
cols = colorRampPalette(rev(brewer.pal(11,"RdBu")))(25)
gcol=brewer.pal(3,"Dark2")
gcol=gcol[sampleInfo$g+1]

##make lables: remove 2005 since it's common to all
labcol= gsub("2005-","",sampleInfo$date)  

##pick highly variable genes:
sds =rowMads(geneExpression)
ind = order(sds,decreasing=TRUE)[1:25]

## make heatmap
heatmap.2(geneExpression[ind,],
          col=cols,
          trace="none",
          scale="row",
          labRow=geneAnnotation$CHR[ind],
          labCol=labcol,
          ColSideColors=gcol,
          key=FALSE)
```

## Question 5
#### A)A:There is no relationship between g and x but with 8,793 tests some will appear significant by chance. Selecting genes with the t-test gives us a deceiving result.
```{r}
set.seed(17)
m = nrow(geneExpression)
n = ncol(geneExpression)
x = matrix(rnorm(m*n),m,n)
g = factor(sampleInfo$g )

library(gplots)
library(matrixStats)
library(genefilter)
library(RColorBrewer)
cols = colorRampPalette(rev(brewer.pal(11,"RdBu")))(25)

ttest = rowttests(x,g)
sds = rowSds(x)
Indexes = list(t=order(ttest$p.value)[1:50], s=order(-sds)[1:50])
for(ind in Indexes){
 heatmap.2(x[ind,],
          col=cols,
          trace="none",
          scale="row",
          labCol=g,
          key=FALSE)
 }

```


# Machine Learning, Exercises on Pages 379-380

## Question 1
```{r}
n = 1000
y = rbinom(n,1,0.25)
##proportion of ones Pr(Y)
sum(y==1)/length(y)

##expectaion of Y
mean(y)

n = 10000
set.seed(1)
men = rnorm(n,176,7) #height in centimeters
women = rnorm(n,162,7) #height in centimeters
y = c(rep(0,n),rep(1,n))
x = round(c(men,women))
##mix it up
ind = sample(seq(along=y))
y = y[ind]
x = x[ind]
```

## Question 2
#### E(Y |X = 176) is equal to 0.1049475
```{r}
mean(y[x==176])
```

## Question 3
#### To guess a female's height with a success probability greater than 0.5, the largest height to predict female would be 168cm.
```{r}
xs = seq(160,178)
pr =sapply(xs,function(x0) mean(y[x==x0]))
plot(xs,pr)
abline(h=0.5)
abline(v=168)

```

# Machine Learning, Exercises on Pages 389-390

## Question 1
#### f(168) is predicted to be 0.5480233
```{r}
 n = 10000
set.seed(1)
men = rnorm(n,176,7) #height in centimeters
women = rnorm(n,162,7) #height in centimeters
y = c(rep(0,n),rep(1,n))
x = round(c(men,women))
##mix it up
ind = sample(seq(along=y))
y = y[ind]
x = x[ind]

set.seed(5)
N = 250
ind = sample(length(y),N)
Y = y[ind]
X = x[ind]

fit=loess(Y~X)
predict(fit,newdata=data.frame(X=168))

##Here is a plot
xs = seq(160,178)
Pr =sapply(xs,function(x0) mean(Y[X==x0]))
plot(xs,Pr)
fitted=predict(fit,newdata=data.frame(X=xs))
lines(xs,fitted)

```

## Question 2
#### The standard error of the loess based estimate is 0.05755689
```{r}
set.seed(5)
B = 1000
N = 250
xs = seq(160,178)
plot(xs,xs,ylim=c(0,1),type="l")
res = replicate(B,{
  ind = sample(length(y),N)
  Y = y[ind]
  X = x[ind]
  fit=loess(Y~X)
  ##optional plots
  fitted=predict(fit,newdata=data.frame(X=xs))
  lines(xs,fitted)
  estimate = predict(fit,newdata=data.frame(X=168))
  return(estimate)
  })

library(rafalib)
popsd(res)
```

```{r}

```

```{r}

```

```{r}

```

```{r}

```


